Optimizing Abstract Abstract Machines

### 2012-11-14 paper/paper.tex
Small fixes.

The final approximation we make to get to our point of departure is
store-allocating results of application sub-expressions. 
The AAM approach stops at the previous optimization. 
However, the $\kfn{}$ continuation stores a value, and this makes the space of
continuations quadratic rather than linear in the size of the program---even for a monovariant analysis like 0CFA.
Having the space of continuations grow linearly with the size of the program 
will drop the overall complexity to cubic (as expected).

@@

The following example illustrates
the essence of abstract compilation effect:
\[
\sapp{\sapp{\sapp\mvar{\mexp_1}}{\mexp_2}}{\mexp_3}
\]
makes the following transitions:
\begin{align}
& \ev{\sapp{\sapp{\sapp\mvar{\mexp_1}}{\mexp_2}}{\mexp_3},\menv,\mkont,\msto_0}\\
\machstep\; &
\ev{\sapp{\sapp\mvar{\mexp_1}}{\mexp_2},\menv,\kar{\mexp_3,\menv,\maddr_1},\msto_1}
\\
\machstep\; &
\ev{\sapp\mvar{\mexp_1},\menv,\kar{\mexp_2,\menv,\maddr_2},\msto_2}
\\
\machstep\; &
\ev{\mvar, \menv,\kar{\mexp_1,\menv,\maddr_3},\msto_3} % {\mexp_2}
\\
\machstep\; &
\co{\kar{\mexp_1,\menv},\mval,\msto_4} % {\mexp_1}{\mexp_2}
\mbox{ where } \mval \in \msto(\menv(\maddr))
\end{align}

where $\msto_4 = \msto_0 \sqcup \{ [\maddr_1 \mapsto \{ \mkont \}],
[\maddr_2 \mapsto \kar{\mexp_3,\menv,\maddr_1}]
[\maddr_3 \mapsto \kar{\mexp_2,\menv,\maddr_2}]$.

@@

\begin{figure}
\begin{align*}
\mathit{eval}(\mexp) &= \{ \mstate\ |\ \compile{\mexp}(\epsilon,\varnothing,\varnothing,\kmt) \multimachstep \mstate \} \text{ where }
\\[2mm]
%% CONTINUE
\co{\kmt,\mval,\msto} &\machstep
\ans{\msto,\mval}
\\
\co{\kar[^\mcntr_\mlab]{\mcomp,\menv,\maddr},\mval,\msto} & \machstep
\mcomp^\mcntr(\menv,\msto,\kfn[^\mcntr_\mlab]{\mval,\maddr})
\\
\co{\kfn[^\mcntr_\mlab]{{\mvalx{u}},\maddr},\mval,\msto} & \machstep
\ap[^\mcntr_\mlab]{\mval,\mvalx{u},\mkont,\msto}
\text{ where }\mkont \in \msto(\maddr)
\\
\co{\kif[^\mcntr]{\mcompi0,\mcompi1,\menv,\maddr},\strue,\msto} & \machstep
\mcompi0^\mcntr(\menv,\msto,\mkont)
\text{ where }\mkont\in\msto(\maddr)
\\
\co{\kif[^\mcntr]{\mcompi0,\mcompi1,\menv,\maddr},\sfalse,\msto} & \machstep
\mcompi1^\mcntr(\menv,\msto,\mkont)
\text{ where }\mkont\in\msto(\maddr)
\\[2mm]
%% APPLY
\ap[^\mcntr_\mlab]{\clos{\mvar,\mcomp,\menv},\mval,\msto,\mkont} & \machstep
\mcomp^{\mcntr'}(\menv',\msto',\mkont) \\
\text{ where }\menv',\msto',\mcntr' &= \mathit{bind\,}^{ \mcntr}_\mlab(\menv,\msto,\mvar,\mval)
\\
\ap{\mop,\mval,\msto,\mkont} & \machstep
\co{\mkont,\mval',\msto} \\
\text{ where }\mkont &\in\msto(\maddr)
\text{ and } \mval'\in\interpdelta(\mop,\mval)
\end{align*}
\caption{Abstract$^2$ machine for compiled ISWIM}
\label{fig:caam}
\end{figure}

@@

\paragraph{Mutable literal data}
The Scheme we model is nonstandard in that it has immutable literal
data (the Scheme benchmarks did not mutate their literal data). To
support mutable literal data, abstract values like {\tt QPair} would
be like regular conses, only point to a preallocated pair of addresses
that hold {\tt QData} as well. This is because conses can be mutated
to hold any value, not just quoted data.

>>>>

The final approximation we make to get to our point of departure is
store-allocating results of application sub-expressions. 
The AAM approach stops at the previous optimization. 
However, the {\tt fn} continuation stores a value, and this makes the space of
continuations quadratic rather than linear in the size of the program---even for a monovariant analysis like 0CFA.
Having the space of continuations grow linearly with the size of the program 
will drop the overall complexity to cubic (as expected).

@@

The following example illustrates
the essence of abstract compilation effect:
\[
\sapp{\sapp{\sapp\mvar{\mexp_1}}{\mexp_2}}{\mexp_3}
\]
makes the following transitions:
\begin{align}
& \ev{\sapp{\sapp{\sapp\mvar{\mexp_1}}{\mexp_2}}{\mexp_3},\menv,\mkont,\msto_0}\\
\machstep\; &
\ev{\sapp{\sapp\mvar{\mexp_1}}{\mexp_2},\menv,\kar{\mexp_3,\menv,\maddr_1},\msto_1}
\\
\machstep\; &
\ev{\sapp\mvar{\mexp_1},\menv,\kar{\mexp_2,\menv,\maddr_2},\msto_2}
\\
\machstep\; &
\ev{\mvar, \menv,\kar{\mexp_1,\menv,\maddr_3},\msto_3} % {\mexp_2}
\\
\machstep\; &
\co{\kar{\mexp_1,\menv},\mval,\msto_4} % {\mexp_1}{\mexp_2}
\mbox{ where } \mval \in \msto(\menv(\maddr))
\end{align}
where \(
\msto_4 = \msto_0 \sqcup \{ [\maddr_1 \mapsto \{ \mkont \}],
[\maddr_2 \mapsto \kar{\mexp_3,\menv,\maddr_1}],
[\maddr_3 \mapsto \kar{\mexp_2,\menv,\maddr_2}]\).

@@

\begin{figure}
\begin{gather*}
\begin{align*}
\mathit{eval}(\mexp) &= \{ \mstate\ |\ \compile{\mexp}(\epsilon,\varnothing,\varnothing,\kmt) \multimachstep \mstate \} \text{ where }
\end{align*}
\\[2mm]
\begin{align*}
%% CONTINUE
\co{\kmt,\mval,\msto} &\machstep
\ans{\msto,\mval}
\\
\co{\kar[^\mcntr_\mlab]{\mcomp,\menv,\maddr},\mval,\msto} & \machstep
\mcomp^\mcntr(\menv,\msto,\kfn[^\mcntr_\mlab]{\mval,\maddr})
\\
\co{\kfn[^\mcntr_\mlab]{{\mvalx{u}},\maddr},\mval,\msto} & \machstep
\ap[^\mcntr_\mlab]{\mval,\mvalx{u},\mkont,\msto}
\text{ where }\mkont \in \msto(\maddr)
\\
\co{\kif[^\mcntr]{\mcompi0,\mcompi1,\menv,\maddr},\strue,\msto} & \machstep
\mcompi0^\mcntr(\menv,\msto,\mkont)
\text{ where }\mkont\in\msto(\maddr)
\\
\co{\kif[^\mcntr]{\mcompi0,\mcompi1,\menv,\maddr},\sfalse,\msto} & \machstep
\mcompi1^\mcntr(\menv,\msto,\mkont)
\text{ where }\mkont\in\msto(\maddr)
\\[2mm]
%% APPLY
\ap[^\mcntr_\mlab]{\clos{\mvar,\mcomp,\menv},\mval,\msto,\mkont} & \machstep
\mcomp^{\mcntr'}(\menv',\msto',\mkont) \\
\text{ where }\menv',\msto',\mcntr' &= \mathit{bind\,}^{ \mcntr}_\mlab(\menv,\msto,\mvar,\mval)
\\
\ap{\mop,\mval,\msto,\mkont} & \machstep
\co{\mkont,\mval',\msto} \\
\text{ where }\mkont &\in\msto(\maddr)
\text{ and } \mval'\in\interpdelta(\mop,\mval)
\end{align*}
\end{gather*}
\caption{Abstract abstract machine for compiled ISWIM}
\label{fig:caam}
\end{figure}

@@
